Clustering Constraints on the Relative Sizes of Central and Satellite Galaxies

### 2017-11-09 paper/galsize_paper.tex
Added comment in conclusions that mocks will be made available upon publication
Finished review of now-complete draft. Paper is ready for co-author review

Observational constraints on models for galaxy $\rhalf$ typically come from one-point function measurements such as $\median{\rhalf}{\mstar}$ and $\sigma(\rhalf\vert\mstar),$ e.g.,  \citet{khochfar_silk06,desmond_etal17,bottrell_etal17b,hou_etal17,somerville_etal17}, or otherwise from catalogs of galaxy groups \citep{weinmann_etal08,huertas_company_etal13b,spindler_wake17}. More recently, connection of galaxy sizes to virial radius of their parent halos have been
explored by converting stellar mass to halo mass and virial radius, using abundance matching \citep{kravtsov_etal04,tasitsiomi_etal04,vale_ostriker04,vale_ostriker06,conroy_etal06}.  Such analysis indicated that sizes of galaxies
are, on average, proportional to the virial radii of their host halos, both at $z=0$ \citep{kravtsov13}, and higher redshifts \citep{huang_etal17,somerville_etal17}.

Observations of two-point galaxy clustering, $\wproj(\rproj),$ have historically been used to place tight constraints on many features of the galaxy--halo connection, such as its $\mstar-$dependence \citep{moster_etal10,leauthaud_etal11,reddick_etal13,skibba_etal15}, dependence on luminosity \citep{kravtsov_etal04, tasitsiomi_etal04,vale_ostriker04,vale_ostriker06,tinker_etal05,cacciato_etal13}, broadband color \citep{coil_etal08,zehavi_etal11,guo_etal11,hearin_watson13}, and star-formation rate \citep{wang_etal07,tinker_etal13,watson_etal14}. In the present work, we exploit the constraining power of $\wproj(\rproj)$ to test empirical models connecting galaxy stellar masses and sizes to the properties of the galaxy's parent dark matter halo.

@@

Using galaxies observed in the Data Release 10 of the Sloan Digital Sky Survey \citep[SDSS,][]{ahn_etal14} with the stellar mass measurements taken from the MPA-JHU catalog \citep{kauffmann_etal03,brinchmann_etal04}, we define volume-limited samples of galaxies according to the same $\mstar-$completeness cuts used in \citet{behroozi_etal15} (see Figure 2). We supplement the MPA-JHU catalog with measurements of half-light radius, $\rhalf,$ derived from galaxy profile decompositions provided by \citet{meert_etal15}. The \citet{meert_etal15} catalog is based on Data Release 7 of SDSS \citep{abazajian_etal09}, with improvements to the photometry pipeline and light profile fitting method and, especially, in background subtraction \citep{vikram_etal10,bernardi_etal13,bernardi_etal14,meert_etal13}. In the version of this catalog that we use,\footnote{Our $\mstar$ and $\rhalf$ measurements are derived from different photometry pipelines; this is driven by our desire for consistency with the \citet{behroozi_etal15} completeness cuts used in our clustering measurements. We note that we have repeated our analysis for stellar mass measurements based on \citet{meert_etal15}, finding only minor quantitative, and no qualitative changes to our results.} two-dimensional $r-$band profiles were fit with a two-component de Vaucouleurs + exponential profile to determine the half-light radius of total $r-$band luminosity, $\rhalf.$

As the bedrock of our modeling, we use the publicly available\footnote{\url{http://www.slac.stanford.edu/~behroozi/BPlanck\_Hlists}} catalog of {\tt Rockstar} subhalos identified at $z=0$ in the Bolshoi-Planck simulation \citep{klypin_etal11,behroozi12_rockstar,behroozi12_consistent_trees,riebe_etal13,rodriguez_puebla16_bolplanck}. As described in \S\ref{subsec:sham}, we will use traditional abundance matching to connect stellar mass $\mstar$ with subhalo peak mass $\mpeak.$ To address issues related to subhalo incompleteness \citep{guo_white13,campbell_etal17}, we supplement the Bolshoi-Planck subhalo catalog with subhalos that have been disrupted and no longer appear in the standard catalog as $z=0$ surviving subhalos. We describe our treatment of these ``orphan" subhalos in the Appendix.

For our SDSS galaxy samples, we calculate two-point clustering $\wproj$ using line-of-sight projection of $\pi_{\rm max}=20~\mpc$ using the {\tt correl} program in {\tt UniverseMachine}. For mock galaxies, to compute galaxy clustering we employ the distant observer approximation by treating the simulation $z-$axis as the line-of-sight. We compute $\wproj$ using the {\tt mock\_observables.wp} function in {\tt Halotools} \citep{hearin_etal16}, which is a python implementation of the algorithm in the {\tt Corrfunc} C library \citep{sinha_etal17}. We also use {\tt Halotools} to compute the galaxy-galaxy lensing signal, $\Delta\Sigma,$ using the {\tt mock\_observables.delta\_sigma} function.

@@

In addition to stellar mass, $\wproj(\rproj)$ depends on galaxy color \citep[e.g.,][]{zehavi_etal11}. Thus, we will also consider
dependence of clustering on galaxy size while controlling for color. The results for the subsamples split in stellar mass and color are presented in \S\ref{subsec:colormorph}.

@@

We map $\mstar$ onto subhalos using deconvolution abundance matching (SHAM). See section 3.3 of \citet{behroozi_etal10} or the appendix of \citet{kravtsov_etal14} for an account of the mathematics underlying this technique in the presence of scatter. To perform the abundance matching, we use the publicly available code\footnote{\url{https://bitbucket.org/yymao/abundancematching/overview}} developed by Yao-Yuan Mao \citep{lehmann_etal15} that provides a python wrapper around the C kernel developed by Peter Behroozi \citep{behroozi_etal10}. For the subhalo property used in SHAM, we use $\mpeak,$ the largest value of $\mvir$ ever attained by the subhalo throughout the entire history of the main progenitor halo.  Using the stellar mass function provided in \citet{moustakas_etal13}, we model $\mstar$ as a log-normal distribution with $0.2$dex of scatter about the median relation $\median{\mstar}{\mpeak}$ determined by SHAM. As described in the appendix, we include a prescription for supplementing the ordinary {\tt Rockstar} subhalo catalog with disrupted subhalos, so that some of our model galaxies inhabit subhalos that are no longer resolved and do not appear in the standard publicly available catalog.

Figure \ref{fig:baseline_sham_clustering} shows that our implementation of SHAM produces model galaxies whose projected clustering, $\wproj(\rproj)$, is in reasonably good agreement with SDSS galaxies. We note, however, that $\wproj(\rproj\gtrsim1{\rm Mpc})$ is $\sim10-20\%$ too weak at stellar mass $\mstar\gtrsim10^{10.75}\msun.$

%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/baseline_sham.pdf}
\caption{
{\bf Abundance matching clustering predictions.}
Using the abundance matching methods described in \S\ref{subsec:sham}, we compare the projected clustering of mock vs. SDSS galaxies. Each panel shows the comparison for a volume-limited sample defined by a different $\mstar-$threshold. Black points with error bars show SDSS measurements; solid black curves show the abundance matching prediction of our fiducial model, which includes the effect of orphan subhalos (see Appendix A). Including orphans mitigates the discrepancy in the clustering predicted by traditional, $\mpeak-$based SHAM based, though mild tension remains on small scales for $\mstar\gtrsim10^{10.75}\msun,$ and on all scales for the $\mstar\gtrsim10^{9.75}\msun.$
}
\label{fig:baseline_sham_clustering}
\end{figure}
%-----------------------------------------------------------------------------------------------------

@@

\textcolor{blue}{AK: I don't think the name of the model is descriptive here, but I could not yet come up with a better one.}
In the first model we explore, we assume that stellar mass $\mstar$ controls $\rhalf,$ so that galaxy sizes are drawn from a log-normal distribution with $0.2$ dex of scatter centered at $\median{\rhalf}{\mstar}.$ To implement this model, for simplicity we tabulate $\median{\rhalf}{\mstar}$ directly from the data, rather than pursue a parametric form \citep[see, e.g.,][]{zhang_yang17}. Note that this model reproduces the observed $\rhalf-\mstar$ relation by construction.

@@

As noted in \S\ref{sec:intro}, \citet{kravtsov13} showed that galaxy sizes are on average proportional to the virial radius of their host halo. Motivated by these results, we  explore a model in which $\rhalf$ scales linearly with the halo virial radius:\footnote{We note that the normalization factor in \citet{kravtsov13} is 0.015, where as our normalization is 0.01. The difference derives from two factors: {\em i.} the choice of halo radius definition: \citet{bryan_norman98} in the present work, compared to the $200c$ definition used in \citet{kravtsov13}; {\em ii.} the latter study used deprojected (larger) sizes for spheroidal galaxies, whereas we make no such correction here.}
\beq
\label{eq:fiducial_model}
\rhalf = 0.01\rvir.
\eeq
For the virial radius of halos and subhalos, we use $\rmpeak:$ the value of $\rvir$ in physical units of $\kpc$ measured at the time when halo mass reached its maximum.  The relationship between $\rmpeak$ and $\mpeak$ is given by
\beq
\mpeak\equiv\frac{4\pi}{3}\rmpeak^{3}\Delta_{\rm vir}(\zpeak)\rho_{\rm m}(\zpeak),
\eeq
where for $\Delta_{\rm vir}(\zpeak)$ we use the fitting function to the ``virial" definition used in \citet{bryan_norman98}. For the model we refer to as the ``$\rvir-$based model", we add uncorrelated log-normal scatter of $\sigmarhalf=0.2$ dex to generate a Monto Carlo realization of the model population.

@@

Figure~\ref{fig:mstarmodelclustering} shows a similar comparison of the observed clustering ratios to the model
in which size is controlled by $\mstar$ with sizes of satellites affected by stellar mass loss due to tides (\S\ref{subsubsec:strippingmodel}). Green solid lines show predictions of the model that does not account for stellar
mass loss of satellites due to tides. In this model $\rhalf$ follows a purely random log-normal distribution centered at $\median{\rhalf}{\mstar}$ with the median relation matched to the observed relation by construction.  This model thus predicts no dependence of galaxy clustering upon $\rhalf$ at fixed $\mstar,$ simply because the scatter of $\rhalf$ about $\median{\rhalf}{\mstar}$ is uncorrelated with any other variable. The figure shows that this model clearly is at odds with significant dependence of clustering on size exhibited by SDSS galaxies.

@@

Figure \ref{fig:colorclustering} shows the clustering of large vs. small galaxies, separately for red and blue samples. For each color-selected sample, large galaxies are slightly more strongly clustered relative to small galaxies. Comparing Figure \ref{fig:colorclustering} to the the upper right panel of Figure \ref{fig:rvir_only_clustering_absolute} shows the size dependence for color-selected samples is dramatically weaker.   For $\mstar-$complete samples, small galaxies cluster much stronger than large galaxies; for color-selected samples, the reverse is true, but the difference is much smaller and is confined to a limited range of scales.  Although the two results may seem surprising, even contradictory, they can, in fact be
explained within a unified simple model, as will be discussed in \S\ref{sec:interpretation}.

%---------------------------------------------------------------------------------------------------
\begin{figure*}
\centering
\includegraphics[width=8cm]{FIGS/color_selected_size_dependent_clustering.pdf}
\caption{
{\bf Distinct $\rhalf-$dependence of clustering for color-selected galaxy samples.}
Projected clustering of SDSS galaxies in the bin of stellar mass: $10^{9.75}<\mstar/\msun<10^{10.25}.$ For both ``blue" and ``red" samples, we first divide SDSS galaxies falling in the $\mstar$ bin based on a color cut at $g-r=0.65.$ We then subsequently compute $\median{\rhalf}{\mstar; {\rm red}}$ and $\median{\rhalf}{\mstar; {\rm blue}}$. In analogy to the left panel of Figure \ref{fig:sizedefinition}, this sample definition results in identical stellar mass function for large and small subsamples, separately for both red and blue parent samples. We then measure $\wproj$ for each of the four subsamples, plotting the results with color-coding as specified in the legend. At this stellar mass, pre-selecting galaxy samples by color results in a dramatic weakening of the $\rhalf-$dependence of $\wproj$ relative to $\mstar-$complete samples, as well as a sign-reversal of the signal.
}
\label{fig:colorclustering}
\end{figure*}
%---------------------------------------------------------------------------------------------------

%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/rvir_only_cen_sat_sizes.pdf}
\caption{
{\bf Relative sizes of centrals and satellites.}
In a narrow bin of halo mass $\mhalo\equiv\mpeak\approx10^{12}\msun,$ we show the distribution of model galaxy sizes for different subpopulations galaxies, as predicted by the $\rvir-$based model. The red histogram shows the sizes of satellites; the blue histogram shows host halos that have never passed inside the virial radius of a larger halo (``true centrals"); the green histogram host halos that were subhalos inside a larger at some point in their past history (``splashback centrals"). In the $\rvir-$based model, galaxy size is set by the {\em physical} size of the virial radius at the time the halo attains its peak mass, naturally resulting in smaller sizes for satellites and splashback centrals relative to true centrals of the same $\mhalo.$
}
\label{fig:censatsizehist}
\end{figure}
%-----------------------------------------------------------------------------------------------------

@@

Recall that in the $\rvir$-based size model, galaxy size is proportional to the virial radius of host halo at the redshift  when the halo halo attained its maximum mass. Thus, we can expect that in this model satellite galaxies will have sizes smaller than sizes of central galaxies hosted by halos of similar stellar mass.  Indeed, Figure \ref{fig:censatsizehist} shows the $\rhalf$ distributions of central, satellite, and ``splashback'' central galaxies hosted by halos of the same mass $\mhalo\equiv\mpeak\approx10^{12}\msun.$ A ``splashback central"  is defined as a present-day central galaxy that used to be a satellite, i.e., its main progenitor halo passed inside the virial radius of a larger halo at some point in its past history. On the other hand, we define a ``true central" as a galaxy that has never been a satellite. The figure shows that there is a clear trend of sizes decreasing from the central to splashback and to satellite galaxies.

@@

Likewise, the significantly weaker dependence of clustering on size for color-selected samples (Figure \ref{fig:colorclustering}), and its reversed sign, can be naturally explained within the same framework.  For example, it is well known that satellites are redder than centrals of the same stellar mass \citep[e.g.,][]{vdB_etal08}; accordingly, red galaxies cluster much more strongly than  blue galaxies \citep[e.g.,][]{zehavi_etal11}. Samples of red (blue) galaxies with stellar masses of $\mstar\sim 10^{10}\ M_\odot$ are dominated by satellite (central) galaxies, and splitting by size within such samples no longer corresponds to the split into satellite- and central-dominated samples. In this case, the $\rvir$-based size model predicts that splitting by size {\em does} correspond to splitting by halo mass $\rhalf\propto\rvir\propto\mhalo^{1/3},$ and therefore that large galaxies should be somewhat more clustered than the small ones. That is, the $\rvir-$based model predicts that $\rhalf-$dependent clustering should be reversed for color-selected samples, and that the magnitude of the dependence should be significantly weaker due to weak dependence of virial radius on halo mass, and weak dependence of clustering on halo mass in this regime.

@@

The success of the $\rvir$-based model in explaining the size dependence of galaxy clustering implies that the sizes of {\it all} galaxies,  red, blue, spheroid, disks, satellites and central, are, on average, proportional to the $\rmpeak,$ the physical size of the virial radius at the time at which $\mhalo$ stops growing. Although correlation is not necessarily causation, such a universal and simple dependence poses a useful challenge to fine-grained galaxy formation models, as it indicates tight connection between the growth of galaxies and their host halos.

The $\rvir$-based size model assumes as an ansatz that galaxy size remains approximately constant after a halo stops growing; as shown in Figure \ref{fig:clustering_ratio_upshot}, this assumption is consistent with the observed size-dependence of galaxy clustering in SDSS.  The halos of typical satellites reach $\mpeak$ far outside the virial radius of their ultimate host halo \citep{behroozi_etal14}, and so in the $\rvir-$based model the size difference between satellites and centrals is largely in place prior to the time of satellite infall. Halos of central and satellite galaxies of the same $\mstar$ differ in the epoch $\zpeak$ at which their halo stopped growing, and this difference gives a natural explanation for their different median sizes.

@@

%The broad features of the $\wproj(\rproj)$ measurements shown in Figure \ref{fig:colorclustering} give insight into the characteristics that should be exhibited by any model for the relationship between $B/T$ and $\rhalf.$ First, the bottom right panel of Figure \ref{fig:colorclustering} indicates that the size of the bulge has only a tenuous connection to the properties of its parent dark matter halo.\footnote{Strictly speaking, the size of the bulge could in principle be closely connected to any number of dark matter halo properties, so long as those halo properties do not significantly impact two-point clustering.} Second, for disk-dominated galaxies, the observed $\rhalf-$dependent clustering suggests that at fixed $\mstar,$ the size of the disk is strongly correlated with whether or not the galaxy is a satellite. These findings imply that bulge prominence is intimately connected with the quenching of star-formation, but that there is wide variation in star-formation histories amongst disk-dominated galaxies, in agreement with numerous previous findings \citep[see, e.g.,][]{woo_etal15,font_etal17}.

% A possibly overkill caveat paragraph:
%While these clustering measurements give supporting evidence that satellite mass stripping plays a sub-dominant role in shaping galaxy size, we point out that our treatment of satellite mass loss is only approximate: we have modified the size of satellites without self-consistently modifying total stellar mass. In the present work, we have opted not to develop and fine-tune a complex, semi-analytic model; instead, we have made a simple ``first-order" estimate for the level at which satellite-specific stripping impacts the two-point function. Our findings support the idea that the small sizes of satellites arise primarily from the satellite's distinctive evolution within a high density region and a steep tidal field {\em prior} to becoming a present-day satellite.

@@

As shown in \citet{campbell_etal15}, group-finding methods are subject to significant systematics, particularly when analyzing trends based on color: contemporary algorithms {\em generically} wash out differences between centrals and satellites; trends can even be artificially created that are not truly present. Our clustering-based, forward modeling methods are immune to such systematics, and so it is encouraging that our findings are commensurable with these results. Once galaxies have been color-selected, Figure \ref{fig:colorclustering} shows that the observed clustering trends nearly vanish in magnitude, and reverse sign. Insofar as blue galaxies are largely disk-dominated, the clustering results in Figure \ref{fig:colorclustering} are also consistent with the disk model presented in \citet{dutton_etal08,dutton_etal10}, which predicts that at fixed stellar mass, there is weak, positive correlation between disk size and halo mass.

@@

We have presented new measurements of the dependence of galaxy clustering upon galaxy size, and used {\tt Halotools} to identify the basic ingredients that influence the signal. We conclude with a brief summary of our primary findings:

>>>>

Observational constraints on models for galaxy $\rhalf$ typically come from one-point function measurements such as $\median{\rhalf}{\mstar}$ and $\sigma(\rhalf\vert\mstar),$ e.g.,  \citet{khochfar_silk06,desmond_etal17,bottrell_etal17b,hou_etal17,somerville_etal17}, or otherwise from catalogs of galaxy groups \citep{weinmann_etal08,huertas_company_etal13b,spindler_wake17}. More recently, the connection between a galaxy's size and the virial radius of its parent halo has been
explored by converting stellar mass to halo mass and virial radius, using abundance matching \citep{kravtsov_etal04,tasitsiomi_etal04,vale_ostriker04,vale_ostriker06,conroy_etal06}.  Such analyses have indicated that galaxy size is, on average, proportional to host halo virial radius, both at $z=0$ \citep{kravtsov13}, and higher redshifts \citep{huang_etal17,somerville_etal17}.

Observations of two-point galaxy clustering, $\wproj(\rproj),$ have historically been used to place tight constraints on many features of the galaxy--halo connection, such as its $\mstar-$dependence \citep{moster_etal10,leauthaud_etal11,reddick_etal13,skibba_etal15}, dependence on luminosity \citep{kravtsov_etal04, tasitsiomi_etal04,vale_ostriker04,vale_ostriker06,tinker_etal05,cacciato_etal13}, broadband color \citep{coil_etal08,zehavi_etal11,guo_etal11,hearin_watson13}, and star-formation rate \citep{wang_etal07,tinker_etal13,watson_etal14}. In the present work, we exploit the constraining power of $\wproj(\rproj)$ to test empirical models connecting galaxy stellar mass and size to the properties of the galaxy's parent dark matter halo.

@@

Using galaxies from Data Release 10 of the Sloan Digital Sky Survey \citep[SDSS,][]{ahn_etal14} with the stellar mass measurements taken from the MPA-JHU catalog \citep{kauffmann_etal03,brinchmann_etal04}, we define volume-limited samples of galaxies according to the same $\mstar-$completeness cuts used in \citet{behroozi_etal15} (see Figure 2). We supplement the MPA-JHU catalog with measurements of half-light radius, $\rhalf,$ derived from galaxy profile decompositions provided by \citet{meert_etal15}. The \citet{meert_etal15} catalog is based on Data Release 7 of SDSS \citep{abazajian_etal09}, with improvements to the photometry pipeline and light profile fitting method and, especially, in background subtraction \citep{vikram_etal10,bernardi_etal13,bernardi_etal14,meert_etal13}. In the version of this catalog that we use,\footnote{Our $\mstar$ and $\rhalf$ measurements are derived from different photometry pipelines; this is driven by our desire for consistency with the \citet{behroozi_etal15} completeness cuts used in our clustering measurements. We note that we have repeated our analysis for stellar mass measurements based on \citet{meert_etal15}, finding only minor quantitative, and no qualitative changes to our results.} two-dimensional $r-$band profiles were fit with a two-component de Vaucouleurs + exponential profile to determine the half-light radius of total $r-$band luminosity, $\rhalf.$

As the bedrock of our modeling, we use the publicly available\footnote{\url{http://www.slac.stanford.edu/~behroozi/BPlanck\_Hlists}} catalog of {\tt Rockstar} subhalos identified at $z=0$ in the Bolshoi-Planck simulation \citep{klypin_etal11,behroozi12_rockstar,riebe_etal13,behroozi_etal12b,rodriguez_puebla16_bolplanck}. As described in \S\ref{subsec:sham}, we use traditional abundance matching to connect stellar mass $\mstar$ with subhalo mass. To address issues related to subhalo incompleteness \citep{guo_white13,campbell_etal17}, we supplement the Bolshoi-Planck subhalo catalog with subhalos that have been disrupted and no longer appear in the standard catalog as $z=0$ surviving subhalos. We describe our treatment of these ``orphan" subhalos in the Appendix.

For our SDSS galaxy samples, we calculate two-point clustering $\wproj$ using line-of-sight projection of $\pi_{\rm max}=20~\mpc$ using the {\tt correl} program in {\tt UniverseMachine}. For mock galaxies, we compute $\wproj$ using the {\tt mock\_observables.wp} function in {\tt Halotools} \citep{hearin_etal16}, which is a python implementation of the algorithm in the {\tt Corrfunc} C library \citep{sinha_etal17}. We also use {\tt Halotools} to compute the galaxy-galaxy lensing signal, $\Delta\Sigma,$ using the {\tt mock\_observables.delta\_sigma} function.

@@

In addition to stellar mass, $\wproj(\rproj)$ depends on galaxy color \citep[e.g.,][]{zehavi_etal11}. Thus, we will also consider
dependence of clustering on galaxy size while controlling for color. Observations of clustering exhibited by subsamples split in stellar mass and color are presented in \S\ref{subsec:colormorph}.

@@

We map $\mstar$ onto subhalos using deconvolution abundance matching (SHAM). See Section 3.3 of \citet{behroozi_etal10}, or the appendix of \citet{kravtsov_etal14}, for an account of the mathematics underlying this technique in the presence of scatter. To perform the abundance matching, we use the publicly available code\footnote{\url{https://bitbucket.org/yymao/abundancematching/overview}} developed by Yao-Yuan Mao \citep{lehmann_etal15} that provides a python wrapper around the C kernel developed by Peter Behroozi \citep{behroozi_etal10}. For the subhalo property used in SHAM, we use $\mpeak,$ the largest value of $\mvir$ ever attained by the subhalo throughout the entire history of the main progenitor halo.  Using the stellar mass function provided in \citet{moustakas_etal13}, we model $\mstar$ as a log-normal distribution with $0.2$dex of scatter about the median relation $\median{\mstar}{\mpeak}$ determined by SHAM. As described in the appendix, we include a prescription for supplementing the ordinary {\tt Rockstar} subhalo catalog with disrupted subhalos, so that some of our model galaxies inhabit subhalos that are no longer resolved and do not appear in the standard publicly available catalog.

Figure \ref{fig:baseline_sham_clustering} shows that our implementation of SHAM produces model galaxies whose projected clustering, $\wproj(\rproj)$, is in reasonably good agreement with SDSS galaxies. We note, however, that $\wproj(\rproj\gtrsim1{\rm Mpc})$ is $\sim10-20\%$ too weak at stellar mass $\mstar\gtrsim10^{10.75}\msun,$ and similarly on all scales for the $\mstar\gtrsim10^{9.75}\msun$ sample.

%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/baseline_sham.pdf}
\caption{
{\bf Abundance matching clustering predictions.}
Using the abundance matching methods described in \S\ref{subsec:sham}, we compare the projected clustering of mock vs. SDSS galaxies. Each panel shows the comparison for a volume-limited sample defined by a different $\mstar-$threshold. Black points with error bars show SDSS measurements; solid black curves show the abundance matching prediction of our fiducial model, which includes the effect of orphan subhalos (see Appendix A). Including orphans mitigates the discrepancy in the clustering predicted by traditional, $\mpeak-$based SHAM, though mild tension remains on small scales for $\mstar\gtrsim10^{10.75}\msun,$ and on all scales for $\mstar\gtrsim10^{9.75}\msun.$
}
\label{fig:baseline_sham_clustering}
\end{figure}
%-----------------------------------------------------------------------------------------------------

@@

\textcolor{red}{AK: I don't think the name of the model is descriptive here, but I could not yet come up with a better one.}
In the first model we explore, we assume that stellar mass $\mstar$ controls $\rhalf,$ so that galaxy sizes are drawn from a log-normal distribution with $0.2$ dex of scatter centered at $\median{\rhalf}{\mstar}.$ To implement this model, for simplicity we tabulate $\median{\rhalf}{\mstar}$ directly from the data, rather than pursue a parametric form \citep[see, e.g.,][]{zhang_yang17}. Thus this model reproduces the observed $\rhalf-\mstar$ relation by construction.

@@

As noted in \S\ref{sec:intro}, \citet{kravtsov13} showed that galaxy sizes are on average proportional to the virial radius of their host halo. Motivated by these results, we  explore a model in which $\rhalf$ scales linearly with halo virial radius:\footnote{We note that the normalization factor in \citet{kravtsov13} is 0.015, where as our normalization is 0.01. The difference derives from two factors: {\em i.} the choice of halo radius definition: \citet{bryan_norman98} in the present work, compared to the $200c$ definition used in \citet{kravtsov13}; {\em ii.} the latter study used deprojected (larger) sizes for spheroidal galaxies, whereas we make no such correction here.}
\beq
\label{eq:fiducial_model}
\rhalf = 0.01\rvir.
\eeq
For the virial radius of halos and subhalos, we use $\rmpeak:$ the value of $\rvir$ in physical units of $\kpc$ measured at the time when halo mass reached its maximum, $\zpeak.$  The relationship between $\rmpeak$ and $\mpeak$ is given by
\beq
\mpeak\equiv\frac{4\pi}{3}\rmpeak^{3}\Delta_{\rm vir}(\zpeak)\rho_{\rm m}(\zpeak),
\eeq
where for $\Delta_{\rm vir}(\zpeak)$ we use the fitting function to the ``virial" definition used in \citet{bryan_norman98}. For the model we refer to as the ``$\rvir-$based model", we add uncorrelated log-normal scatter of $\sigmarhalf=0.2$ dex to generate a Monto Carlo realization of the model population.

@@

Figure~\ref{fig:mstarmodelclustering} shows a similar comparison of the observed clustering ratios to the model
in which size is controlled by $\mstar$ with sizes of satellites affected by stellar mass loss due to tides (\S\ref{subsubsec:strippingmodel}). Green solid lines show predictions of the model that does not account for stellar
mass loss of satellites due to tides. In this model $\rhalf$ follows a purely random log-normal distribution centered at $\median{\rhalf}{\mstar}$ with the median relation matched to the observed relation by construction.  This model thus predicts no dependence of galaxy clustering upon $\rhalf$ at fixed $\mstar,$ simply because the scatter of $\rhalf$ about $\median{\rhalf}{\mstar}$ is uncorrelated with any other variable. The figure shows that this model clearly is at odds with the significant dependence of clustering on size exhibited by SDSS galaxies.

@@

Figure \ref{fig:colorclustering} shows the clustering of large vs. small galaxies, separately for red and blue samples. For each color-selected sample, large galaxies are slightly more strongly clustered relative to small galaxies. Comparing Figure \ref{fig:colorclustering} to the the upper right panel of Figure \ref{fig:rvir_only_clustering_absolute} shows the size dependence for color-selected samples is dramatically weaker.   For $\mstar-$complete samples, small galaxies cluster much stronger than large galaxies; for color-selected samples, the reverse is true, but the difference is much smaller and is confined to a limited range of scales.  Although the two results may seem surprising, even contradictory, they can, in fact be explained within a simple, unified model, as will be discussed in \S\ref{sec:interpretation}.

%---------------------------------------------------------------------------------------------------
\begin{figure*}
\centering
\includegraphics[width=8cm]{FIGS/color_selected_size_dependent_clustering.pdf}
\caption{
{\bf Distinct $\rhalf-$dependence of clustering for color-selected galaxy samples.}
Projected clustering of SDSS galaxies in the bin of stellar mass: $10^{9.75}<\mstar/\msun<10^{10.25}.$ For both ``blue" and ``red" samples, we first divide SDSS galaxies falling in the $\mstar$ bin based on a color cut at $g-r=0.65.$ We then subsequently compute $\median{\rhalf}{\mstar; {\rm red}}$ and $\median{\rhalf}{\mstar; {\rm blue}}$. In analogy to the left panel of Figure \ref{fig:sizedefinition}, this sample definition results in identical stellar mass function for large and small subsamples, separately for both red and blue parent samples. We then measure $\wproj$ for each of the four subsamples, plotting the results with color-coding as specified in the legend. At this stellar mass, pre-selecting galaxy samples by color results in a dramatic weakening of the $\rhalf-$dependence of $\wproj$ relative to $\mstar-$complete samples, as well as a sign-reversal of the signal. In \S\ref{sec:interpretation} we give a simple halo model picture unifying the trends shown here with those illustrated in Figure \ref{fig:clustering_ratio_upshot}. 
}
\label{fig:colorclustering}
\end{figure*}
%---------------------------------------------------------------------------------------------------

%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/rvir_only_cen_sat_sizes.pdf}
\caption{
{\bf Relative sizes of centrals and satellites.}
In a narrow bin of halo mass $\mhalo\equiv\mpeak\approx10^{12}\msun,$ we show the distribution of model galaxy sizes for different subpopulations galaxies, as predicted by the $\rvir-$based model. The red histogram shows the sizes of satellites; the blue histogram shows host halos that have never passed inside the virial radius of a larger halo (``true centrals"); the green histogram host halos that were subhalos inside a larger halo at some point in their past history (``splashback centrals"). In the $\rvir-$based model, galaxy size is set by the {\em physical} size of the virial radius at the time the halo attains its peak mass, naturally resulting in smaller sizes for satellites and splashback centrals relative to true centrals of the same $\mhalo.$
}
\label{fig:censatsizehist}
\end{figure}
%-----------------------------------------------------------------------------------------------------

@@

Recall that in the $\rvir$-based size model, galaxy size is proportional to halo virial radius at $\zpeak,$ the redshift when the halo attained its maximum mass. Thus we can expect that in this model, satellite galaxies will have smaller sizes than  central galaxies hosted by halos of similar mass.  Indeed, Figure \ref{fig:censatsizehist} shows the $\rhalf$ distributions of central, satellite, and ``splashback'' central galaxies hosted by halos of the same mass, $\mhalo\equiv\mpeak\approx10^{12}\msun.$ A ``splashback central"  is defined as a present-day central galaxy that used to be a satellite, i.e., its main progenitor halo passed inside the virial radius of a larger halo at some point in its past history. On the other hand, we define a ``true central" as a galaxy that has never been a satellite. The figure shows that there is a clear trend of sizes decreasing from the central to splashback and to satellite galaxies.

@@

Likewise, the significantly weaker dependence of clustering on size for color-selected samples (Figure \ref{fig:colorclustering}), and its reversed sign, can be naturally explained within the same framework.  For example, it is well known that satellites are redder than centrals of the same stellar mass \citep[e.g.,][]{vdB_etal08}; accordingly, red galaxies cluster much more strongly than  blue galaxies \citep[e.g.,][]{zehavi_etal11}. Samples of red (blue) galaxies with stellar masses of $\mstar\sim 10^{10}\ M_\odot$ are dominated by satellite (central) galaxies, and splitting by size {\em within} such samples no longer corresponds to the split into satellite- and central-dominated samples. In this case, the $\rvir$-based size model predicts that splitting by size {\em does} correspond to splitting by halo mass $\rhalf\propto\rvir\propto\mhalo^{1/3},$ and therefore that large galaxies should be somewhat more clustered than the small ones. That is, the $\rvir-$based model predicts that $\rhalf-$dependent clustering should be reversed for color-selected samples, and that the magnitude of the dependence should be significantly weaker due to weak dependence of virial radius on halo mass, and weak dependence of clustering on halo mass in this regime.

@@

The success of the $\rvir$-based model in explaining the size dependence of galaxy clustering implies that the sizes of {\it all} galaxies,  red, blue, spheroid, disks, satellites and central, are, on average, proportional to the $\rmpeak,$ the physical size of the virial radius at the time at which $\mhalo$ stops growing. Although correlation is not necessarily causation, such a universal and simple dependence poses a useful challenge to fine-grained galaxy formation models, as it indicates a tight connection between the growth of galaxies and their host halos.

The $\rvir$-based size model assumes as an ansatz that galaxy size remains approximately constant after its halo stops growing; as shown in Figure \ref{fig:clustering_ratio_upshot}, this assumption is consistent with the observed size-dependence of galaxy clustering in SDSS.  The halos of typical satellites reach $\mpeak$ far outside the virial radius of their ultimate host halo \citep{behroozi_etal14}, and so in the $\rvir-$based model the size difference between satellites and centrals is largely in place prior to the time of satellite infall. Halos of central and satellite galaxies of the same $\mstar$ differ in the epoch $\zpeak$ at which their halo stopped growing, and this difference gives a natural explanation for their different median sizes.

@@

@@

As shown in \citet{campbell_etal15}, group-finding methods are subject to significant systematics, particularly when analyzing trends based on color: contemporary algorithms {\em generically} wash out differences between centrals and satellites; trends can even be artificially created that are not truly present. Our forward modeling inference technique is not subject to such systematics because our methods are entirely outwith the group-finding framework, and so it is encouraging that our findings are commensurable with these results. Once galaxies have been color-selected, Figure \ref{fig:colorclustering} shows that the observed clustering trends nearly vanish in magnitude, and reverse sign. Insofar as blue galaxies are largely disk-dominated, our clustering results in Figure \ref{fig:colorclustering} are also consistent with the disk model presented in \citet{dutton_etal08,dutton_etal10}, which predicts that at fixed stellar mass, there is weak, positive correlation between disk size and halo mass. Our conclusions are also qualitatively consistent with \citet{spindler_wake17}, who find that centrals in the \citet{yang_etal05a} group catalog are $\sim20-25\%$ larger than satellites of similar halo mass (using stellar velocity dispersion as an observational proxy for halo mass); we find the same trend at a comparable, if slightly larger magnitude.

@@

We have presented new measurements of the dependence of galaxy clustering upon galaxy size, and used forward-modeling methods based on {\tt Halotools} to identify the basic ingredients that influence the signal. We conclude with a brief summary of our primary findings:
